---
title: "Homework 6 zz2901"
author: "Troy Zhou"
output: github_document
---

```{r load_package, echo = FALSE, message=FALSE}
library(tidyverse)
library(purrr)
options(tibble.print_min = 5)
options(pillar.sigfig = 5)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

```

## Problem 1

## Problem 2

Create a city_state variable (e.g. “Baltimore, MD”), and a binary variable indicating whether the homicide is solved. Omit cities Dallas, TX; Phoenix, AZ; and Kansas City, MO – these don’t report victim race. Also omit Tulsa, AL – this is a data entry mistake. For this problem, limit your analysis those for whom victim_race is white or black. Be sure that victim_age is numeric.

```{r homicide_new_var}

homicide_clean <- read_csv("data/homicide-data.csv") %>% 
  mutate(resolved = as.numeric(disposition == "Closed by arrest"),
         victim_age = as.numeric(recode(victim_age,"Unknown"="")),
         city_state = paste0(city, ", ", state)) %>% 
  filter(!(city_state %in% c("Dallas, TX","Phoenix, AZ","Kansas City, MO","Tulsa, AL"))
         & victim_race %in% c("White", "Black"))

homicide_clean
```

For the city of Baltimore, MD, use the glm function to fit a logistic regression with resolved vs unresolved as the outcome and victim age, sex and race as predictors. Save the output of glm as an R object; apply the broom::tidy to this object; and obtain the estimate and confidence interval of the adjusted odds ratio for solving homicides comparing male victims to female victims keeping all other variables fixed.

```{r}
fit_logistic = 
  homicide_clean %>% 
  filter(city_state=="Baltimore, MD") %>% 
  glm(resolved ~ victim_age + victim_sex + victim_race, data = ., 
      family = binomial())

baltimore_df = fit_logistic %>% 
  broom::tidy() %>% 
  mutate(OR = exp(estimate),
         LCL = exp(estimate-1.96*std.error),
         UCL = exp(estimate+1.96*std.error)) %>%
  select(term, log_OR = estimate, OR, LCL, UCL) %>% 
  knitr::kable(digits = 3)

baltimore_df
```

* From the output in the table, the estimated odds ratio of having homicide solved between male and female victims is 0.426, adjusting for victim age and race, in Baltimore, MD. 

* We are 95% confident that the true odds ratio of having homicide solved between male and female victims lies between 0.325 and 0.558, adjusting for victim age and race, in Baltimore, MD. 

Now run glm for each of the cities in your dataset, and extract the adjusted odds ratio (and CI) for solving homicides comparing male victims to female victims. Do this within a “tidy” pipeline, making use of purrr::map, list columns, and unnest as necessary to create a dataframe with estimated ORs and CIs for each city.

```{r}
calc_or <- function(x){

  glm(resolved ~ victim_age + victim_sex + victim_race, data = x,
      family = binomial()) %>% 
  broom::tidy() %>%
  mutate(OR = exp(estimate),
         LCL = exp(estimate-1.96*std.error),
         UCL = exp(estimate+1.96*std.error)) %>%
  filter(term=="victim_sexMale") %>% 
  select(log_OR = estimate, OR, LCL, UCL)

}


homicide_nest = homicide_clean %>% 
  nest(data = -city_state) %>% 
  mutate(estimates = map(data,calc_or)) %>% 
  unnest(estimates) %>% 
  select(city_state, OR, LCL, UCL)

homicide_nest

```


Create a plot that shows the estimated ORs and CIs for each city. Organize cities according to estimated OR, and comment on the plot.

```{r}
homicide_nest %>% 
  ggplot(aes(reorder(city_state,OR), OR)) +       
  geom_point(aes(color=city_state)) +
  geom_errorbar(aes(ymin = LCL, ymax = UCL))+
    labs(
    title = "Problem 2, plot of Adjusted OR and CI for solving homicides comparing male victims to female victims In each city ",
    x = "City, State",
    y = "Adjusted OR") +
   theme(axis.text.x = element_text(angle = 90),
         legend.position = "none")
```

* From the plot, we can see the majority of adjusted ORs have their values and upper CI below 1, which means mostly male victims have smaller odds of solved homicide case than females.

* Decent amount of ORs are not significent as their CIs contain the null value of 1. 

## Problem 3

Load and clean the data for regression analysis (i.e. convert numeric to factor where appropriate, check for missing data, etc.).

```{r}
birth_wt <- read.csv("data/birthweight.csv") %>% 
  mutate(babysex=as.factor(babysex),
         frace=as.factor(frace),
         mrace=as.factor(mrace),
         malform=as.factor(malform),
         id=row_number())

str(birth_wt)

sapply(birth_wt, function(x) sum(is.na(x)))
```

* `babysex`, `frace`, `mrace` were converted to factor variables. 

* No missing values in the dataframe.

Propose a regression model for birthweight. This model may be based on a hypothesized structure for the factors that underly birthweight, on a data-driven model-building process, or a combination of the two. Describe your modeling process and show a plot of model residuals against fitted values – use add_predictions and add_residuals in making this plot.

```{r}
library(modelr)
library(mgcv)

set.seed(1)

train_df = sample_n(birth_wt, 800)
test_df = anti_join(birth_wt, train_df, by = "id")


birth_wt %>% 
  ggplot(aes(x = wtgain, y = bwt)) + 
  geom_point(alpha=.5) + 
  geom_smooth()+
  labs(title = "Problem 3, mother’s weight gain during pregnancy (pounds) vs. birth weight",
       x = "mother’s weight gain during pregnancy (pounds)",
       y = "birth weight (grams)"
  )

birth_wt %>% 
  ggplot(aes(x=smoken,y = bwt)) + 
  geom_point(alpha=.5) + 
  geom_smooth() + 
  labs(title = "Problem 3, number of cigarettes smoked daily vs. birth weight",
       x = "number of cigarettes smoked daily",
       y = "birth weight (grams)"
  )

birth_wt %>% 
  ggplot(aes(x=malform,y = bwt)) + 
  geom_boxplot(aes(group=malform)) + 
  labs(title = "Problem 3, boxplots of birth weight stratified by presence of malformations",
       x = "presence of malformations",
       y = "birth weight (grams)"
  )

birth_wt %>% 
  ggplot(aes(x = momage, y = bwt)) + 
  geom_point(alpha=.5) + 
  geom_smooth() + 
  labs(title = "Problem 3, mother’s age at delivery (years) vs. birth weight",
       x = "mother’s age at delivery (years)",
       y = "birth weight (grams)"
  )

birth_wt %>% 
  ggplot(aes(x = delwt, y = bwt)) + 
  geom_point(alpha=.5) + 
  geom_smooth() + 
  labs(title = "Problem 3, mother’s weight at delivery (pounds) vs. birth weight",
       x = "mother’s weight at delivery (pounds)",
       y = "birth weight (grams)"
  )
```

* From exploring the distribution of some suspected variables versus birth weight, we can see no clear relationship between cigarette smoked and birth weight nor between presence of malformations and birth weight nor between mother's age at delivery and birth weight.

* However, some correlation between mom's weight gain and birth weight as well as between weight at delivery and birth weight.

* Therefore, we will use weight gain and weight at delivery as the predictors. 

```{r}

linear_mod = lm(bwt ~ wtgain+delwt, data = train_df)
smooth_mod = gam(bwt ~ s(wtgain)+s(delwt), data = train_df)
wiggly_mod = gam(bwt ~ s(wtgain)+s(delwt, k = 30), sp = 10e-6, data = train_df)


rmse(linear_mod, test_df)
rmse(smooth_mod, test_df)
rmse(wiggly_mod, test_df)
```

* From the three graphs, we can tell the smooth or standard gam fit reflect a more accurate diminishing positive relationship.

* Combined with the fact that the smooth fit has the smallest root mean squared errors, the standard gam model might be better.

```{r}
train_df %>% 
  add_predictions(smooth_mod) %>% 
  add_residuals(smooth_mod) %>% 
  ggplot(aes(x = pred, y = resid)) + geom_point() + 
  geom_smooth( color = "red") +
  labs(title = "Problem 3, plot for residuals against fitted values in training df",
       x = "residual",
       y = "fitted value"
  )
```

* From the plot we can see residuals are fairly evenly distributed around 0 in training df. 


Compare your model to two others:

1) One using length at birth and gestational age as predictors (main effects only)

2) One using head circumference, length, sex, and all interactions (including the three-way interaction) between these

Make this comparison in terms of the cross-validated prediction error; use crossv_mc and functions in purrr as appropriate.

Note that although we expect your model to be reasonable, model building itself is not a main idea of the course and we don’t necessarily expect your model to be “optimal”.

```{r}

  crossv_mc(birth_wt, 100) %>% 
  mutate(
    train = map(train, as_tibble),
    test = map(test, as_tibble),
    current_mod = map(train, ~gam(bwt ~ s(wtgain) + s(delwt), data = .x)),
    other_mod_1  = map(train, ~lm(bwt ~ blength + gaweeks, data = .x)),
    other_mod_2  = map(train, ~lm(bwt ~ bhead + blength + babysex + bhead*blength + bhead*babysex + bhead*blength*babysex, data = .x))) %>% 
  mutate(
    rmse_current = map2_dbl(current_mod, test, ~rmse(model = .x, data = .y)),
    rmse_mod_1 = map2_dbl(other_mod_1, test, ~rmse(model = .x, data = .y)),
    rmse_mod_2 = map2_dbl(other_mod_2, test, ~rmse(model = .x, data = .y))) %>% 
  select(starts_with("rmse")) %>% 
  pivot_longer(
    everything(),
    names_to = "model", 
    values_to = "rmse",
    names_prefix = "rmse_") %>% 
  mutate(model = fct_inorder(model)) %>% 
  ggplot(aes(x = model, y = rmse)) + geom_violin()

```

* The model chosen before using weight gain during pregnancy and weight at delivery is clearly "worse" than the other two given its much higher RMSE.

* However, this was expected as the other two models are predicting baby weight using the baby's measurements (head circumference and baby length). These predictors are highly correlated with birth weight but carry little clinical meaning. There is not much point in predicting a baby's birthright after they are born and using their measurements.  

* Using my proposed predictors of mothers' weight measurements, we can at least actually predict the baby's weight before birth and study what factors are actually affecting birth weight. Hence, the proposed model is more clinically meansingful. 








